Interpretable Risk Minimization -- Preliminaries

### 2024-08-05 paper/main.tex
updated model weights, plots and paper score report
explained D_KL modification on writeup

We therefore derive an optimization objective that attempts to maximize the contrast between the two activations maps. For the binary classification case, $(c, c')$ reflects that correct and incorrect class respectively.
\begin{gather}
	\max_{\theta} s_c - s_{c'} \equiv \max_{\theta} \tilde{\mathcal{\bm{A}}}_{c}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c'}^{\text{HiResCAM}}
\end{gather}
This is equivalent to the original ERM optimization objective, with the difference being the resultant values, which is an activation map instead. If $\tilde{\mathcal{\bm{A}}}_{c}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c'}^{\text{HiResCAM}} > 0$, the classification is correct -- if not, the classification is incorrect.

@@

Segmentations maps provide this exact information. It describes the region over which the object is present. We therefore define a constrastive cost function as follows:
\begin{gather}
	\{(x, (h, y))\}^N \in \mathcal{D} \\
	\mathcal{\bm{A}}^{\text{contrastive}} := \tilde{\mathcal{\bm{A}}}_{y}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{y'}^{\text{HiResCAM}} \\
	\mathcal{\bm{A}}^{\text{background}}_{i,j} := \begin{cases}
		\mathcal{\bm{A}}^{\text{contrastive}}_{i,j} & h_{i,j} = 0 \\
		0 & \text{otherwise}
	\end{cases} \\
	\mathcal{\bm{A}}^{\text{foreground}}_{i,j} := \begin{cases}
		0 & h_{i,j} = 0 \\
		\mathcal{\bm{A}}^{\text{contrastive}}_{i,j} & \text{otherwise}
	\end{cases} \\
	\mathcal{\bm{J}}(\theta, \mathcal{D}) := D_{KL}(\sigma_{\text{softmax}}(\lambda_1 \mathcal{\bm{A}}^{\text{contrastive}})\ ||\ \sigma_{\text{softmax}}(\lambda_2 \cdot h)) + \frac{\lambda_3}{D_1 D_2} ||\mathcal{\bm{A}}^{\text{background}}||^2_2 - \frac{1}{D_1 D_2} \sum^{D_1,D_2}_{i,j} \mathcal{\bm{A}}^{\text{foreground}}_{i,j}
\end{gather}

@@

@@

\begin{table}[h]
	\centering
	\begin{tabular}{c|ccc}
		\toprule
		\textbf{Method}  & \textbf{Cross Entropy Loss}  & \textbf{Training Accuracy}   & \textbf{Validation Accuracy} \\
		\midrule
		Cross-Entropy & 0.628 & 68.2\% & 64.1\% \\
		Interpretable (Ours) & \bf 0.354 & \bf 96.2\% & \bf 89.2\% \\
		\bottomrule
	\end{tabular}
\end{table}

>>>>

We therefore derive an optimization objective that attempts to maximize the contrast between the two activations maps. For the binary classification case, $(c, c')$ reflects that correct and incorrect class respectively.
\begin{gather}
	\max_{\theta} s_c + b_c - s_{c'} - b_{c'} \equiv \max_{\theta} \tilde{\mathcal{\bm{A}}}_{c}^{\text{HiResCAM}} + b_c - \tilde{\mathcal{\bm{A}}}_{c'}^{\text{HiResCAM}} - b_{c'}
\end{gather}
This is equivalent to the original ERM optimization objective, with the difference being the resultant values, which is an activation map instead. If $\tilde{\mathcal{\bm{A}}}_{c}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c'}^{\text{HiResCAM}} > 0$, the classification is correct -- if not, the classification is incorrect. Ignoring the bias terms, our new optimization objective is:
\begin{gather}
	\max_{\theta} \tilde{\mathcal{\bm{A}}}_{c}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c'}^{\text{HiResCAM}}
\end{gather}

@@

Segmentations maps provide this exact information. It describes the region over which the object is present. We therefore define a constrastive cost function as follows:
\begin{gather}
	\{(x, (h, y))\}^N \in \mathcal{D} \\
	\mathcal{\bm{A}}^{\text{contrastive}} := \tilde{\mathcal{\bm{A}}}_{y}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{y'}^{\text{HiResCAM}} \\
	\mathcal{\bm{A}}^{\text{background}}_{i,j} := \begin{cases}
		\mathcal{\bm{A}}^{\text{contrastive}}_{i,j} & h_{i,j} = 0 \\
		0 & \text{otherwise}
	\end{cases} \\
	\mathcal{\bm{A}}^{\text{foreground}}_{i,j} := \begin{cases}
		0 & h_{i,j} = 0 \\
		\mathcal{\bm{A}}^{\text{contrastive}}_{i,j} & \text{otherwise}
	\end{cases} \\
	D_{KL}(P || Q) := P \cdot \log \frac{P'}{Q} \text{ where }
	P'_{i,j} := \begin{cases}
		1 & P_{i,j} \neq 0 \\
		P_{i,j} & \text{otherwise}
	\end{cases} \\
	\mathcal{\bm{J}}(\theta, \mathcal{D}) := D_{KL}(\sigma_{\text{softmax}}(\lambda_1 \mathcal{\bm{A}}^{\text{contrastive}})\ ||\ \sigma_{\text{softmax}}(\lambda_2 \cdot h)) + \frac{\lambda_3}{D_1 D_2} ||\mathcal{\bm{A}}^{\text{background}}||^2_2 - \frac{1}{D_1 D_2} \sum^{D_1,D_2}_{i,j} \mathcal{\bm{A}}^{\text{foreground}}_{i,j}
\end{gather}

@@

$D_{KL}$ is modified such that the likelihood comparator sets point probability to 1 if it is part of the foreground. While this violates the probability distribution definition, it motivates the optimizer to maximize the foreground activation. During training, observing pointwise divergence revealed that the model was penalized for `over-activating' the foreground, which this modification negates.

@@

\begin{table}[h]
	\centering
	\begin{tabular}{c|ccc}
		\toprule
		\textbf{Method}  & \textbf{Cross Entropy Loss}  & \textbf{Training Accuracy}   & \textbf{Validation Accuracy} \\
		\midrule
		Cross-Entropy & 0.628 & 68.2\% & 64.1\% \\
		Interpretable (Ours) & \bf 0.327 & \bf 98.4\% & \bf 90.0\% \\
		\bottomrule
	\end{tabular}
\end{table}
